The First Level of Super Mario Bros.~is Easy with Lexicographic Orderings and Time Travel {\normalsize \ldots after that it gets a little tricky.}

### 2013-03-17 trunk/tasbot/paper/paper.tex
Almost done. Maybe one more game, abstract, layout polish

{\tt ReweightMotifs} changes the weight of motifs that match the best
{\tt next} sequence that we just committed to, if we think that we
made progress on this step. The idea is to learn which sequences tend
to work well for us; this might be different from what the human
player did. For example, in run-to-the-right-and-jump kinds of games,
human players tend to hesitate more before obstacles and enemies than
a computer does.\footnote{That said, this part was an early idea and
  is probably not necessary. It's suspicious because these 10-frame
  sequences are not necessarily motifs (10 is the same as the normal
  motif length, and futures are generated by motifs, but they can
  become desynchronized because of the final incomplete motif, future
  mutation, and other things). So sometimes the chosen sequence
  doesn't affect weights. I think this would be better if we kept a
  Markov model } Knowing whether we made progress is kind of
difficult, since we can almost always find some move that increases
the objectives locally. For this I use the same approach of {\em value
  fraction} from Section~\ref{sec:objectiveweighting} based on a
sample of memories that we've seen so far. If we appear to be
progressing then the motif is weighted up; if we're regressing then it
is weighted down.

@@

\paragraph{Input models.}

\paragraph{Better backtracking.}
XXX Keep backtracking until improvability is below a certain limit.
XXX backtrack farther in some conditions.

@@

XXX

>>>>

{\tt ReweightMotifs} changes the weight of motifs that match the best
{\tt next} sequence that we just committed to, if we think that we
made progress on this step. The idea is to learn which sequences tend
to work well for us; this might be different from what the human
player did. For example, in run-to-the-right-and-jump kinds of games,
human players tend to hesitate more before obstacles and enemies than
a computer does.\footnote{That said, this part was an early idea and
  is probably not necessary. It's suspicious because these 10-frame
  sequences are not necessarily motifs (10 is the same as the normal
  motif length, and futures are generated by motifs, but they can
  become desynchronized because of the final incomplete motif, future
  mutation, and other things). So sometimes the chosen sequence
  doesn't affect weights. I think this would be better if we kept a
  Markov model and updated it no matter what sequences we generated.}
Knowing whether we made progress is kind of difficult, since we can
almost always find some move that increases the objectives locally.
For this I use the same approach of {\em value fraction} from
Section~\ref{sec:objectiveweighting} based on a sample of memories
that we've seen so far. If we appear to be progressing then the motif
is weighted up; if we're regressing then it is weighted down.

@@

\paragraph{Input models.}
I'm unsatisfied with the motif approach. As mentioned earlier, the
obvious thing to try instead is a Markov model, which would probably
be simpler and would allow re-weighting from inputs regardless of what
we concocted while playing (the current version can only reweight the
human motifs, not learn new sequences discovered while running). I
would also like some solution to the start button problem---if it is
among the motifs, it often shows up in gameplay in annoying ways. I
don't feel good about simply blacklisting it, however. In Bubble
Bobble, start appears to be used to burn away futures when all of them
are bad. Maybe a simple improvement would be to allow the inner loop
of {\tt playfun} to skip the turn (empty {\tt next} sequence) in
order to simulate this.

\paragraph{Better backtracking.}
Backtracking is a powerful idea, and there's lots more to do here. The
fixed-size backtracking window is disappointing since it's a
parameter, and doesn't even seem right for most games. It would make
sense to do something like lengthen the window until the {\em
  improvability} starts dropping; basically, find the amount of recent
past that's comparable to random play, and try improving that.
Moreover, it should be possible to re-backtrack, since some parts of
the game are harder than others. Ideally, when Mario has so few
options that he contemplates suicide, he should be backtracking over
and over and farther and farther until he finds a future that's good.
This is the power of time travel. Use it, man.

\paragraph{Efficiency in search}
Quality is directly related to the number of states searched. Some
sequences are easier to search than others, because they share a prefix
with something we've already done and can be cached. It would be worth
looking into algorithms that explicitly take into account the cost of
branching, so that we explore some tree (of futures, for example) rather
than disjoint linear futures. The effective number of futures explored
could be much higher for the same CPU.

@@

In this paper I showed how lexicographic orderings and time travel can
be used to automate classic video games. The approach uses an
amusingly simple and mathematically elegant model to learn what
constitutes ``winning'' according to a human player's inputs. It then
uses hundreds of CPU hours to search different input sequences that
seem to ``win'', inspecting only the RAM of the simulated game,
ignoring any of the human outputs like video and sound. The technique
works great on some games, getting farther into the game than the
human's inputs did, and produces novel gameplay (for example, bug
exploitation and super-human daredevil timing). On other games, it is
garbage, but on a scale of recycling symbol 1 to recycling symbol 7,
it is at least hilarious garbage.
